A sheaf--function dictionary for algebraic tori over local fields

### 2013-08-17 shanghai.tex

\begin{definition}\label{def:GN}
For each positive integer $d$, set $\TT_{d} \ceq \TT \times_\Spec{\OK} \Spec{\OK/\pK^d}$.
Observe that $\OK/\pK^d$ is an Artin ring and $\TT_{d}$ is smooth and locally of finite type over $\OK/\pK^d$.
%
Let $\GN{T}_{d} = \Gr(\TT_d)$ be the group scheme over $\Fq$ produced by applying
the Greenberg functor to $\TT_{d}$; this is the \emph{Greenberg transform of $\TT_d$}.
\end{definition}

See \cite{greenberg:63a} and \cite{bosch-lutkebohmert-reynaud:NeronModels}*{Ch. 9, \S 6}
and the first few sections of \cite{stasinski:12a} for background on the Greenberg functor.

For each positive integer $d$, consider the canonical surjective ring homomorphisms
$\OK/\pK^{d+1} \to \OK/\pK^d$ and the associated cartesian diagram of schemes 
\[
\begin{tikzcd}
\Spec{\OK/\pK^d} \arrow[hook]{rr} && \Spec{\OK/\pK^{d+1}} \arrow[hook]{r} & \Spec{\OK} \\
 \arrow{u} \TT_{d} \arrow[hook]{rr} & \ \arrow[Rightarrow]{d}{\text{Greenberg functor}} & \arrow{u} \TT_{d+1}  &  \\
\GN{T}_{d} \arrow[hook]{rr} & \ & \GN{T}_{d+1}  \arrow{r} & \Spec{k}
\end{tikzcd}
\]
The map $\GN{T}_{d} \hookrightarrow \GN{T}_{d+1}$ is not a morphism of group schemes.
There is, however, an epimorphism of group schemes $\GN{T}_{d+1} \to \GN{T}_{d}$ for
which $\GN{T}_{d} \hookrightarrow \GN{T}_{d+1}$ is a section.

The fundamental properties of $\GN{T}_d$ that we need are:
 \begin{enumerate}
   \labitem{(GN.0)}{dFq} $\GN{T}_d(\Fq) = \TT(\OK/\pK^d)$, canonically.
  \labitem{(GN.1)}{GNd} $\GN{T}_d$ is a smooth, commutative group scheme,
  locally of finite type over $\Fq$, and the component group of $\GN{T}_d$ is
  geometrically finitely generated. (In fact, $\pi_0(\GN{T}_d) = \pi_0(\TT) \times_{\Spec{\OK}} \Spec{\Fq}$,
  and the latter is finitely generated by \cite{xarles:93a}*{2.18}.)
  \labitem{(GN.2)}{d+1} 
  \end{enumerate}

>>>>

We will also need the Greenberg functor, 
introduced in \cite{greenberg:61} and \cite{greenberg:63a}, 
as a functor from the category of schemes locally of finite type 
over an Artinian local ring with perfect residue field $k$ 
to the category of schemes locally of finite type over $k$. 
This functor respects open immersions, closed immersions, 
etale morphisms and smooth morphisms. 
Besides Greenberg's original papers, 
see \cite{bosch-lutkebohmert-reynaud:NeronModels}*{Ch. 9, \S 6} 
and the first few sections of \cite{stasinski:12a} 
for background on the Greenberg functor. 
Let
\[
%\begin{tikzcd}
\Gr_d : \left( \text{Schemes} \atop \text{lft over $\OK$}\right) \to \left( \text{Schemes} \atop \text{lft over $\Fq$}\right)
%\end{tikzcd}
\]
be the functor produced by precomposing pull-back along $\OK \to \OK/\pK^d$ with the Greenberg functor.
The surjective ring homomorphism $\OK/\pK^{d+1} \to \OK/\pK^d$ determines a natural transformation $P_{d} : \Gr_{d+1} \to \Gr_d$. 
%\cite{bosch-lutkebohmert-reynaud:NeronModels}*{p. 276}.

For any positive integer $d$ set $\GN{T}_d = \Gr_d(\TT)$ and $f_d = P_d(\TT)$.
The fundamental properties of $f_d : \GN{T}_{d+1} \to \GN{T}_d$ that we need are:
 \begin{enumerate}
   \labitem{(GN.0)}{dFq} $\GN{T}_d(\Fq) = \TT(\OK/\pK^d)$, canonically.
  \labitem{(GN.1)}{GNd} $\GN{T}_d$ is a smooth, commutative group scheme,
  locally of finite type over $\Fq$, and the component group of $\GN{T}_d$ is
  geometrically finitely generated. (In fact, $\pi_0(\GN{T}_d) = \pi_0(\TT) \times_{\Spec{\OK}} \Spec{\Fq}$,
  and the latter is finitely generated by \cite{xarles:93a}*{2.18}.)
  \labitem{(GN.2)}{truncation}
  The morphism $f_d : \GN{T}_{d+1} \to \GN{T}_d$ is a surjective morphism of group schemes.
  \end{enumerate}

We remark that $f_d : \GN{T}_{d+1} \to \GN{T}_d$ is not induced from a morphism of Néron models.
